Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-22 appendix-EPQ.tex
modify

\EPQRulesAndSemantics*

@@

\EPQClosedUnderProjection*

@@

\DataDifferentiatedisEnoughforPQ*

@@

\ConCheckEPQIsCorrect*

@@

\EPQasMultiInMRpriforHistory*

@@

\priExecutionIsEnough*
\begin {proof}

@@

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$. $\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, and there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $x$ in $e'$.

@@

\EPQOneEqualAsPBandGP*

@@

\OBOrderHasBoundedLength*

@@

An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}. Here we rename the items that covers $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning items into $e$. In this figure, $c = \textit{call}(\textit{put},e,\textit{anyPri}),\textit{ret}(\textit{put},e)$, $\textit{call}(\textit{rm},e), \textit{ret}(\textit{rm},e),\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty})$, $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$, $c_2 = c_1 + \textit{ret}(\textit{put},b)$, $c_3 = c_2 + \textit{call}(\textit{put},d),\textit{ret}(\textit{rm},d)$, $c_4 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{rm},d)$.

@@

An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}. In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $c_1$, $c_2$, $c_3$ and $c_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},b)$ and before $\textit{call}(\textit{rm},a)$: There are two possible positions of $\textit{call}(\textit{put},a,p)$: (1) before $\textit{call}(\textit{rm},a_1)$, and (2) after $\textit{call}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a)$.

@@

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{put},a_1,\_)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a)$ and before $\textit{call}(\textit{put},a_1,p)$: same as above case.

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{call}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.

\item[-] If $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},a_1)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a)$, (2) after $\textit{call}(\textit{put},b,p)$ and before $\textit{call}(\textit{rm},a_1)$, and (3) before $\textit{call}(\textit{put},b,p)$.
\end{itemize}

Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}. For each case, we construct an finite automaton. Let $\textit{Auts}_{\textit{1-eq}}^{3}$ be the set of finite automata that is constructed for above fourteen cases. For example, for the case $\textit{ca}_1$ when $\textit{call}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},a_1)$, $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}. In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $c$ and $c_1 = c + \textit{call}(\textit{put},d,\textit{les}_p)$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}. Let $c_2 = c_1 + \textit{ret}(\textit{put},a_1)$, $c_3 = c_2 + \textit{ret}(\textit{put},b)$, $c_4 = c_3 + \textit{call}(\textit{put},d) + \textit{ret}(\textit{rm},d)$, and $c_5 = c + \textit{ret}(\textit{put},b) + \textit{ret}(\textit{put},a_1) + \textit{ret}(\textit{rm},d)$. The other register automata in $\textit{Auts}_{\textit{1-eq}}^{3}$ can be similarly constructed.

@@

Similarly, we construct sets $\textit{Auts}_{\textit{1-eq}}^{4}$ and $\textit{Auts}_{\textit{1-eq}}^{5}$ of register automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.

Let $\textit{Auts}_{\textit{1-eq}} = \{ \mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2 \} \cup \textit{Auts}_{\textit{1-eq}}^{3} \cup \textit{Auts}_{\textit{1-eq}}^{4} \cup \textit{Auts}_{\textit{1-eq}}^{5}$. The following lemma states that $\textit{PQ}_1^{=}$ is co-regular.

\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
\label{lemma:EPQ1Equal is co-regular}
$\mathsf{MatchedMaxPriority}^{=}$ is co-regular.
\end{restatable}

@@

We need to prove that, given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$, if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e),$ $\seqPQ_1^{=} \in \textit{last}(e') \wedge e'$ does not linearizable w.r.t. $\textit{MS}(\seqPQ_1^{=})$.

@@

\noindent {\bf $\textit{fact}_1$}: Given a data-independence implementation $\mathcal{I}$, $\textit{Auts}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if $\exists e \in \mathcal{I}_{\neq},$ $e' \in \textit{proj}(e)$, $\textit{last}(e')=\seqPQ_1^{=}$, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.

\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\textit{Auts}_{\textit{1-eq}}$. By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$. Since $e_1$ is accepted by some register automata in  $\textit{Auts}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.

let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$. It is obvious that $e'' \in \textit{proj}(e_2)$ is a $p$-execution, $\textit{last}(e'') = \seqPQ_1^{=}$. According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e''$.

\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $last(e')=\textit{PQ}_1^{=}$, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $e'$ is a $p$-execution, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$. By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $e$. It is easy to see that $\textit{last}(e_1) = \seqPQ_1^{=}$, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$. By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists). Then

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $h_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{3}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{4}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $h_1$ is accepted by some register automaton in $\textit{Auts}_{\textit{1-eq}}^{5}$.
\end{itemize}

@@

\subsection{Co-Regular of $\seqPQ_2^{>}$}
\label{subsec:appendix co-regular of EPQ2Lar}

\begin{restatable}{lemma}{EPQ2LarIsAlwaysCoRegular}
\label{lemma:EPQ2Lar is always co-regular}
Given a data-differentiated $\_$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{>}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{>})$.
\end{restatable}

\begin {proof}
Since $\textit{last}(e) = \seqPQ_2^{>}$, the actions with maximal priority in $e$ is some unmatched $\textit{put}$. Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the guard of $\textit{MS}(\seqPQ_2^{>})$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{Co-Regular of $\seqPQ_2^{=}$}
\label{subsec:appendix co-regular of EPQ2Equal}

\begin{restatable}{lemma}{EPQ2EqualAsHappenBefore}
\label{lemma:EPQ2Equal as happen before}
Given a data-differentiated $p$-execution $e$ with $\textit{last}(e) = \seqPQ_2^{=}$. $e$ does not linearizable to $\textit{MS}(\seqPQ_2^{=})$, if and only if there exists $x$ and $y$ with priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
\end{restatable}

@@

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,p)$ does not happen before $\textit{put}(y,p)$. We need to prove that $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.

Let $x_1,\ldots,x_m$ be the set of items with priority $p$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of items with priority $p$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$. By assumption, we know that $\textit{call}(\textit{put},y_i,p)$ is before $\textit{ret}(\textit{put},x_j)$ for each $i,j$. Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:

@@

Let $l$ be the sequence of linearization points. It is easy to see that $e \sqsubseteq l$. Since linearization points of $\textit{put}(x_i,p)$ is after the linearization point of $\textit{put}(y_j,p)$ for each $i,j$, it is easy to see that $l \in \textit{MS}(\seqPQ_2^{=})$. This completes the proof of this lemma. \qed
\end {proof}

Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check violation to $\textit{MS}(\seqPQ_2^{=})$. However, the case in Lemma \ref{lemma:EPQ2Equal as happen before} violates our assumption that each single-priority execution is FIFO. Therefore, we know that $\seqPQ_2^{=}$ is always co-regular, as states by the following lemma.

\begin{restatable}{lemma}{EPQ2EqualIsAlwaysCoRegular}
\label{lemma:EPQ2Equal is always co-regular}
Given a data-differentiated $p$-execution $e$, if $\textit{last}(e) = \seqPQ_2^{=}$, then $e \sqsubseteq \textit{MS}(\seqPQ_2^{=})$.
\end{restatable}

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the items of $e$ or $o$, to which we add a node,

>>>>

%\EPQRulesAndSemantics*

$\newline$

{\noindent \bf Lemma \ref{lemma:EPQ rules and semantics}}: $\textit{Check-PQ-Seq}(e)=\mathsf{true}$ iff $e\in \seqPQ$, for every data-differentiated sequential execution $e$.

@@

%\EPQClosedUnderProjection*

{\noindent \bf Lemma \ref{lem:closure_proj}:} $\seqPQ$ is closed under projection, i.e., $\textit{proj}(e)\subseteq \seqPQ$ for each $e\in \seqPQ$.

@@

%\DataDifferentiatedisEnoughforPQ*

{\noindent \bf Lemma \ref{lemma:data differentiated is enough for PQ}:} A data-independent implementation $\mathcal{I}$ is linearizable w.r.t a data-independent set $S$ of sequential executions, if and only if $\mathcal{I}_{\neq}$ is linearizable w.r.t. $S_{\neq}$.

@@

%\ConCheckEPQIsCorrect*

{\noindent \bf Lemma \ref{lemma:con-check-EPQ is correct}}: $\textit{Check-PQ-Conc}(e)=\mathsf{true}$ iff $e \sqsubseteq \seqPQ$, for every data-differentiated execution $e$.

@@

{\noindent \bf Lemma \ref{lemma:EPQ as multi in MRpri for history}}: Given a data-differentiated execution $e$, $e \sqsubseteq \seqPQ$ if and only if for each $e' \in \textit{proj}(e)$, $\textit{Check-PQ-Conc-NonRec}(e')$ returns $\mathsf{true}$.

%\EPQasMultiInMRpriforHistory*

@@

$\newline$

{\noindent \bf Lemma \ref{lemma:pri execution is enough}}: Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ a data-differentiated execution. Then, $e$ is $\Gamma$-linearizable iff $e\vert_{\preceq p}$ is $\Gamma$-linearizable for some maximal priority $p$ in $e$.

%\priExecutionIsEnough*
\begin {proof}

@@

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$. $\mathcal{A}_{\textit{1-lar}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, and there is a cycle going through $x$ in $G$, where $G$ is the left-right constraint of $x$ in $e'$.

@@

%\EPQOneEqualAsPBandGP*

$\newline$
{\noindent \bf Lemma \ref{lemma:EPQ1Equal as pb order and gap-point}}: Let $e$ be a data-differentiated execution which contains only one maximal priority $p$ such that $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds.
%Given a data-differentiated $p$-execution $e$ where $\mathsf{MatchedMaxPriority}^{=}\mathsf{\text{-}Condition}(e)$ holds.
Then, $e$ is not linearizable w.r.t $\mathsf{MatchedMaxPriority}^{=}$ iff $e$ contains two values $x$ and $y$ of maximal priority $p$ such that $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is strictly smaller than the index of both $\textit{call}(\textit{put},y,p)$ and $\textit{call}(\textit{rm},y)$.

@@

%\OBOrderHasBoundedLength*

$\newline$
{\noindent \bf Lemma \ref{lemma:ob order has bounded length}}: Let $e$ be a data-differentiated execution such that $a <_{\textit{pb}} a_1 <_{\textit{pb}} \ldots <_{\textit{pb}} a_m <_{\textit{pb}} b$ holds for some set of values $a$, $a_1$,$\ldots$,$a_m$, $b$. Then, one of the following holds:
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $a <_{\textit{pb}}^A b$, $a <_{\textit{pb}}^B b$, or $a <_{\textit{pb}}^C b$,

\item[-] $a <_{\textit{pb}}^A a_i <_{\textit{pb}}^B b$ or $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A b$, for some $i$.
%\item[-] $a <_{\textit{pb}}^B a_i <_{\textit{pb}}^A a_j <_{\textit{pb}}^B b$, for some $i$ and $j$,
\end{itemize}

@@

An automaton $\mathcal{A}_{\textit{l-eq}}^1$ is given in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}, and it is constructed for the first enumeration in \figurename~\ref{fig:history enumeration 1 for PQ1Equal}. Here we rename the values that cover $\textit{call}(\textit{rm},a)$ and $\textit{ret}(\textit{rm},b)$ into $d$, and rename the remanning values into $\top$. In this figure, $C = \{ \textit{call}(\textit{put},\top,\textit{true}),\textit{ret}(\textit{put},\top,\textit{true})$, $\textit{call}(\textit{rm},\top), \textit{ret}(\textit{rm},\top),$ $\textit{call}(\textit{rm},\textit{empty}),\textit{ret}(\textit{rm},\textit{empty}) \}$, $C_1 = C \cup \{ \textit{call}(\textit{put},d,<r) \}$, $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_3 = C_2 \cup \{ \textit{call}(\textit{put},d,<r),\textit{ret}(\textit{rm},d) \}$, $C_4 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{rm},d) \}$.

@@

An automaton $\mathcal{A}_{\textit{l-eq}}^2$ is given in \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, and it is constructed for the second enumeration in \figurename~\ref{fig:history enumeration 2 for PQ1Equal}. In \figurename~\ref{fig:automata for second enumeration of PQ1Equal}, $C_1$, $C_2$, $C_3$ and $C_4$ is same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}.

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},b)$ and before $\textit{call}(\textit{rm},a)$: There are two possible positions of $\textit{call}(\textit{put},a,p)$: (1) before $\textit{call}(\textit{rm},a_1)$, and (2) after $\textit{call}(\textit{rm},a_1)$, and before $\textit{ret}(\textit{put},a,p)$.

@@

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{put},a_1,p)$ and before $\textit{ret}(\textit{rm},a_1)$: same as above case.

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{put},a,p)$ and before $\textit{call}(\textit{put},a_1,p)$: same as above case.

\item[-] If $\textit{call}(\textit{put},b,p)$ is after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a,p)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{put},b,p)$ and before $\textit{ret}(\textit{put},a,p)$, (2) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{call}(\textit{put},b,p)$, and (3) before $\textit{call}(\textit{rm},a_1)$.

\item[-] If $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},a_1)$: There are three possible positions of $\textit{call}(\textit{put},a,p)$: (1) after $\textit{call}(\textit{rm},a_1)$ and before $\textit{ret}(\textit{put},a,p)$, (2) after $\textit{call}(\textit{put},b,p)$ and before $\textit{call}(\textit{rm},a_1)$, and (3) before $\textit{call}(\textit{put},b,p)$.
\end{itemize}

Therefore, there are fourteen possible cases that satisfy the third enumeration in \figurename~\ref{fig:history enumeration 3 for PQ1Equal}. For each case, we construct an finite automaton. Let $\mathcal{A}_{\textit{1-eq}}^{3}$ be the union of register automata that is constructed for above fourteen cases. For example, for the case $\textit{ca}_1$ when $\textit{call}(\textit{put},a,p)$ is before $\textit{call}(\textit{rm},a_1)$, $\textit{call}(\textit{put},b,p)$ is after $\textit{ret}(\textit{rm},a_1)$, and $\textit{call}(\textit{put},b,p)$ is before $\textit{call}(\textit{rm},b)$, we construct a finite automaton $\mathcal{A}_{\textit{l-eq}}^{\textit{3-1}}$ in \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}. In \figurename~\ref{fig:automata for ca1 of third enumeration of Rpr2}, let $C$ and $C_1$ the same as that in \figurename~\ref{fig:automata for first enumeration of PQ1Equal}. Let $C_2 = C_1 \cup \{ \textit{ret}(\textit{put},a_1,=r) \}$, $C_3 = C_2 \cup \{ \textit{ret}(\textit{put},b,=r) \}$, $C_4 = C_3 \cup \{ \textit{call}(\textit{put},d,<r), \textit{ret}(\textit{rm},d) \}$, and $C_5 = C \cup \{ \textit{ret}(\textit{put},b,=r), \textit{ret}(\textit{put},a_1,=r), \textit{ret}(\textit{rm},d) \}$. Other register automata can be similarly constructed.

@@

Similarly, we construct sets $\mathcal{A}_{\textit{1-eq}}^{4}$ and $\mathcal{A}_{\textit{1-eq}}^{5}$ of union of register automata for the forth enumeration in \figurename~\ref{fig:history enumeration 4 for PQ1Equal} and the fifth enumeration in \figurename~\ref{fig:history enumeration 5 for PQ1Equal}, respectively.

Let $\mathcal{A}_{\textit{1-eq}}$ be the union of $\mathcal{A}_{\textit{l-eq}}^1, \mathcal{A}_{\textit{l-eq}}^2,\mathcal{A}_{\textit{1-eq}}^{3}, \mathcal{A}_{\textit{1-eq}}^{4}$ and $\mathcal{A}_{\textit{1-eq}}^{5}$. The following lemma states that $\mathcal{A}_{\textit{1-eq}}$ is $\mathsf{MatchedMaxPriority}^{=}$-complete.

\begin{restatable}{lemma}{EPQOneEqualIsCoRegular}
\label{lemma:EPQ1Equal is co-regular}
$\mathcal{A}_{\textit{1-eq}}$ is $\mathsf{MatchedMaxPriority}^{=}$-complete.
\end{restatable}

@@

We need to prove that, given a data-independent implementation $\mathcal{I}$. $\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$ such that $e'$ is not $\mathsf{MatchedMaxPriority}^{=}$-linearizable.

@@

\noindent {\bf $\textit{fact}_1$}: Given a data-independent implementation $\mathcal{I}$. $\mathcal{A}_{\textit{1-eq}} \cap \mathcal{I} \neq \emptyset$ if and only if there exists $e \in \mathcal{I}$ and $e' \in \textit{proj}(e)$, $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds, $x$ is the value with maximal priority in $e'$, $e'$ has only one maximal priority, $a$ and $b$ are two items with maximal priority $p$ in $e'$, $a <_{\textit{pb}}^* b$ in $e'$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e'$.

\noindent The $\textit{only if}$ direction: Assume that $e_1 \in \mathcal{I}$ is accepted by some register automata in $\mathcal{A}_{\textit{1-eq}}$. By data-independence, there exists data-differentiated execution $e_2 \in \mathcal{I}$ and a renaming function $r$, such that $e_1=r(e_2)$. Since $e_1$ is accepted by some register automata in  $\mathcal{A}_{\textit{1-eq}}$, let $x$, $y$ and $z$ (if exists) be the items that are renamed into $b$, $a$ and $a_1$ (if exists) by $r$, respectively, and let $d_1,\ldots,d_m$ be the items that are renamed into $d$ by $r$.

let $e'' = e_2 \vert_{ \{ x,y,z,d_1,\ldots,d_m \} }$. It is obvious that $e'' \in \textit{proj}(e_2)$, $e''$ has only one maximal priority, and $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e'')$ holds. According to our construction of automata in $\textit{Auts}_{\textit{1-eq}}$, it is not hard to see that $x$ and $y$ has maximal priority in $h_2$, $y <_{\textit{pb}}^* x$, and the rightmost gap-point of $x$ is before $\textit{call}(\textit{put},y,p)$ or $\textit{call}(\textit{rm},y)$ in $e''$.

\noindent The $\textit{if}$ direction: Assume that there exists $e \in \mathcal{I}_{\neq},e' \in \textit{proj}(e)$, such that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e')$ holds $e'$ has only one maximal priority, $a'$ and $b'$ are two items with maximal priority $p$ in $e'$, $a' <_{\textit{pb}}^* b'$ in $e'$, and the rightmost gap-point of $b'$ is before $\textit{call}(\textit{put},a',p)$ or $\textit{call}(\textit{rm},a')$ in $e'$. By data-independence, we can obtain execution $e_1$ as follows: (1) rename $a'$ and $b'$ into $a$ and $b$, respectively, (2) for the items $d_1,\ldots,d_m$ that constitute the rightmost gap of $b'$, we rename them into $d$, (3) if $a' <_{\textit{pb}}^A a'_1 <_{\textit{pb}}^B b$, we rename $a'_1$ into $a_1$, and (4) rename the other items into $\top$. It is easy to see that $\mathsf{Has\text{-}MatchedMaxPriority}^{=}(e_1)$ holds, $a$ and $b$ has maximal priority in $e_1$, $a <_{\textit{pb}}^* b$ in $e_1$, and the rightmost gap-point of $b$ is before $\textit{call}(\textit{put},a,p)$ or $\textit{call}(\textit{rm},a)$ in $e_1$. By Lemma \ref{lemma:five enumeration is enough for EPQ1Equal}, there are five possible enumeration of operations of $a$, $b$, $a_1$ (if exists). Then

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $a <_{\textit{pb}}^* b$ because of the first enumeration, it is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^1$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the second enumeration, it is easy to see that $e_1$ is accepted by $\mathcal{A}_{\textit{l-eq}}^2$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the third enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{3}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the forth enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{4}$.

\item[-] If $a <_{\textit{pb}}^* b$ because of the fifth enumeration, it is easy to see that $e_1$ is accepted by some register automaton in $\mathcal{A}_{\textit{1-eq}}^{5}$.
\end{itemize}

@@

\subsection{$\mathsf{UnmatchedMaxPriority}^{>}(e)$ Always Holds}
\label{subsec:appendix co-regular of EPQ2Lar}

\begin{restatable}{lemma}{EPQ2LarIsAlwaysCoRegular}
\label{lemma:EPQ2Lar is always co-regular}

Given a data-differentiated execution $e$ with only one maximal priority, if $\mathsf{Has\text{-}}\mathsf{Unmat-}$ $\mathsf{chedMaxPriority}^{>}(e)$ holds, then $e \sqsubseteq l$ for some $l$ where $\mathsf{UnmatchedMaxPriority}^{>}\mathsf{\text{-}Seq}(l,x)$ holds. 
\end{restatable}

\begin {proof}
Since $\mathsf{Has\text{-}}\mathsf{UnmatchedMaxPriority}^{>}(e)$ holds, the actions with maximal priority in $e$ is some unmatched $\textit{put}$. Therefore, no matter how we locate linearization points, we can always obtain a sequence $l$ of operations that contains unmatched $\textit{put}$ with maximal priority, and this satisfy the requirements of $\mathsf{UnmatchedMaxPriority}^{>}$. This completes the proof of this lemma. \qed
\end {proof}

\subsection{$\mathsf{UnmatchedMaxPriority}^{=}(e)$ Always Holds}
\label{subsec:appendix co-regular of EPQ2Equal}

\begin{restatable}{lemma}{EPQ2EqualAsHappenBefore}
\label{lemma:EPQ2Equal as happen before}
Given a data-differentiated execution $e$ with only one priority and $\mathsf{Has\text{-}}\mathsf{Unmatched-}$ $\mathsf{MaxPriority}^{=}(e)$ holds. $e$ is not $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable, if and only if there exists $x$ and $y$ with maximal priority $p$, $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, and $\textit{put}(x,p) <_{\textit{hb}} \textit{put}(y,p)$.
\end{restatable}

@@

To prove the $\textit{only if}$ direction, we prove its contrapositive. Assume that for each pair of $x$ and $y$ with maximal priority in $e$, if $x$ has unmatched $\textit{put}$, $y$ has matched $\textit{put}$ and $\textit{rm}$, then $\textit{put}(x,p)$ does not happen before $\textit{put}(y,p)$. We need to prove that $e$ is $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable.

Let $x_1,\ldots,x_m$ be the set of values with priority $p$ and has unmatched $\textit{put}$ in $e$, let $y_1,\ldots,y_n$ be the set of values with priority $p$ and has matched $\textit{put}$ and $\textit{rm}$ in $e$. By assumption, we know that $\textit{call}(\textit{put},y_i,p)$ is before $\textit{ret}(\textit{put},x_j,p)$ for each $i,j$. Then we explicitly construction the linearization of $e$ by locating the linearization points of $e$ as follows:

@@

Let $l$ be the sequence of linearization points. It is easy to see that $e \sqsubseteq l$. Since linearization points of $\textit{put}(x_i,p)$ is after the linearization point of $\textit{put}(y_j,p)$ for each $i,j$, it is easy to see that $\mathsf{UnmatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(l,x')$ holds where $x' = x_{\_}$ and its return action is after other values in $x_1,\ldots,x_m$. This completes the proof of this lemma. \qed
\end {proof}

Lemma \ref{lemma:EPQ2Equal as happen before} shows how to check non $\mathsf{UnmatchedMaxPriority}^{=}(e)$-linearizable exections. However, the case in Lemma \ref{lemma:EPQ2Equal as happen before} violates our assumption that each single-priority execution is FIFO. Therefore, we know that $\mathsf{UnmatchedMaxPriority}^{=}(e)$ always holds, as states by the following lemma.

\begin{restatable}{lemma}{EPQ2EqualIsAlwaysCoRegular}
\label{lemma:EPQ2Equal is always co-regular}
Given a data-differentiated execution $e$ with only one maximal priority, if $\mathsf{Has\text{-}}\mathsf{Unmat-}$ $\mathsf{chedMaxPriority}^{=}(e)$ holds, then $e \sqsubseteq l$ for some $l$ where $\mathsf{UnmatchedMaxPriority}^{=}\mathsf{\text{-}Seq}(l,x)$ holds. 
\end{restatable}

@@

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] the nodes are the values of $e$ or $o$, to which we add a node,

\item[-] there is an edge from value $d_1$ to $o$, if $\textit{put}(d_1,\_)$ happens before $o$,
